---
title: "Data Visualization and Modelling"
output:
  pdf_document: default
  html_notebook: default
---

# 1. Simulation

In statistics, is usual that people are interested in generate values of uniformly distributed random variables. In fact, many pocket calculators have a function that provides _random numbers_. Even programming languages such as _C_ or _Visual Basic_ have also a function that generates uniformly distributed random variables $$U(0,1)$$.

**How does the computer or the pocket calculator generate independent** $$U(0,1)$$ *values?*

These values are generated from completely deterministic algorithms. The _LCG (Linear Congruential Generator)_ algorithms are the most frequent.

For example: $$U_n = frac(9821*U_{n-1} + 0.211327$$, where _frac_ indicates the _fractionary part_ and $$U_0$$ is the _seed_. For instance, if we set $$U_0 = 0.1$$, then
```{r}
frac <- function(x){
	y <- x - trunc(x)
	return(y)
}

n <- 5
U <- numeric(n)
seed <- 0.1

for(i in 1:n){
	U[i] <- frac(9821 * seed + 0.211327)
	seed <- U[i]
}

print(U)

```

However, we must keep in  mind that these are _pseudo-random_ numbers, because they are generated by using **deterministic** algorithms.

> _"Anyone who considers arithmetical methods of producing random digits is, of course, in a state of sin"._, John Von neumann, 1951

## Does randomness even exist?

> "We may regard the present state of the universe as the effect of its past and the cause of its  future. An intellect which at any given moment knew all of the forces that animate nature and the mutual positions of the beings that compose it, if this intellect  were vast  enough to submit the data to analysis, could condense into a single formula the movement of the greatest bodies of the universe and that of the lightest atom; for such an intellect nothing could be uncertain and the future just like the past would be present before its eyes." _Pierre Simon de Laplace_

One could say that given  knowledge of the initial  state (position, speed, etc.), classical physics procedures can be used to perfectly predict the outcome of a coin toss. Thus, _randomness_ is a consequence of our knowledge. Notice how this argument can be treacherous, for no one has ever constructed a predictive model for a coin toss.

Actually, nowadays we have better tools for generating random numbers. For instance the _Quantum Random Number Generators_, which using semi-plated mirror and photonics theory provides a 0-1 independent Bernoulli $p=1/2$ series. 

This can be really useful in terms of generating uniform distributed random numbers. For instance, consider a number in  basis 2:

```{r}
number <- c(1,0,0,1,1,0,0,1)
print(number)
```
This series correspons to the number 

```{r}
rnumber <- c()

for(i in 1:length(number)){
	rnumber[i] <- number[i]/(2^i)
}
fnumber <- sum(rnumber)
print(fnumber)
```

# 2. Classical Generic  Methods

Now that we know how to generate uniformly distributed random numbers, we can think about how can other random variables could be generated.

### Exaxmple: Simulating a dice using $U(0,1)$

```{r}
p <- 1/6
d <- numeric(1/p)
for(j in 1:10000){
	for(i in 1:(1/p)){
		u <- runif(1)
		if((i-1)*p<u & u<i*p){d[i] <- d[i]+1}
	}
}
library(ggplot2)
ggplot() +
	geom_bar(data=d)
```

